---
title: "Javanese Names and Gender Equality Progress"
author: "Firman M Firmansyah"
date: "March 18, 2021"
output: html_document
---

These codes are supplementary materials for the following manuscript.

APA
Firmansyah, F. M., Pratama, A. R., & Wijayanto, A. W. (2021, October 7). Naming Practice and Gender Equality Progress in Javanese Society: A Computational Approach. https://doi.org/10.31235/osf.io/dacu7

IEEE
F. M. Firmansyah, A. R. Pratamaand A. W. Wijayanto, “Naming Practice and Gender Equality Progress in Javanese Society: A Computational Approach”, 07-Oct-2021. [Online]. Available: osf.io/preprints/socarxiv/dacu7.

# PREPARATION
## Required Library

```{r, results='hide', message=FALSE, echo=FALSE, warning=FALSE}

library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library(summarytools)
library(rstatix)

```

## Dataset
### Attach original Central Java dataset
This data set is needed to perform: IDENTIFYING, CLEANING, and COUNTING. 

```{r}

central_java_df <- read_rds(file.choose()) #original data set scrapped from KPU's website: jateng.rds
central_java_df$ID <- 1:nrow(central_java_df) #create new ID

```

### Attach/save counted dataset
This data set results from IDENTIFYING, CLEANING, and COUNTING.

```{r}

#saveRDS(javanese_native_df, file = "javanese_native_df.rds") #non running

#attach dataset
javanese_native_df <- readRDS("javanese_native_df.rds")

```

### Save Shared dataset
This data set is anonymous, consisting of native (place of birth: urban, suburb, rural), year of birth (YOB), gender, name length, and character length. 

```{r}

anon_javanese_native_df <- javanese_native_df[,c(3:5,8:9)]

#saveRDS(anon_javanese_native_df, file = "anon_javanese_native_df.rds")

```


## Functions
### identify_pob
This function maps place of birth (POB) to: urban, suburb, rural, and migrant.

```{r}

identify_pob <- function(x) {
  if (x %in% cj_urban) {
    y <- "urban"
  }
  else if (x %in% cj_suburb) {
    y <- "suburb"
  }
  else if (x %in% cj_rural) {
    y <- "rural"
  }
  else {
    y <- "migrant"
  }
  return(y)
}

```

### nwords
This function counts numbers of words in a name (as words).

```{r}

nwords <- function(x) {
  y <- str_count(x, "\\S+")
  return(y)
}

```

### ncharacters
This function counts numbers of characters in a name (as characters).

```{r}

ncharacters <- function(x) {
  y <- str_replace_all(x, pattern=" ", repl="")
  z <- nchar(y)
  return(z)
}

```

### stn_err
This function calculates standard errors (se).
It is needed to perform DESCRIPTIVE STATS.

```{r}

stn_err <- function(x) {
  y <- sd(x)/sqrt(length(x))
  return(y)
}

```


# IDENTIFYING
## Identifying place of birth (POB)
Only those who were born and voted in Central Java are deemed native Javanese.
People who were born in Special Region of Yogyakarta or East Java, even though may be Javanese too, are excluded from the analysis.

```{r}

javanese_df <- tibble(ID=central_java_df$ID, POB=central_java_df$TL) # create a new tibble, with place of birth

javanese_df$POB <- str_replace_all(javanese_df$POB, "[^[:alnum:]]", " ") # clear place of birth
javanese_df$POB <- str_replace_all(javanese_df$POB, pattern=" ", repl="") # remove space

# only include place of birth with N >= 50
place_of_birth <- javanese_df %>%
  group_by(POB) %>%
  summarise(N = length(ID)) %>%
  arrange(desc(N), POB) %>%
  filter(N >= 50)

place_of_birth

# list of urban, suburb, and rural areas in Central Java

cj_urban <- c("tegal", "magelang", "semarang", "pekalongan", "surakarta", "salatiga", "solo", "kotasemarang", "kotamagelang", "magelamg") 
              
cj_suburb <- c("brebes", "brebbes", "kabbrebes", "klaten", "kabklaten", "kabsemarang", "sukoharjo", "purwokerto", "kabtegal", "purwodadi", "ambarawa", "ungaran", "kabkaranganyar", "gombong", "kutoarjo", "cepu", "kabmagelang", "bumiayu", "karangamyar", "kabupatentegal", "muntilan", "kabupatensemarang", "karangayar", "kabsmg", "kendal", "demak", "karayanganyar", "batang", "kabupatenkendal", "kabbatang", "kabkendal", "kabdemak") 
                             
cj_rural <- c("cilacap", "banyumas", "kebumen", "grobogan", "pemalang", "pati", "banjarnegara", "jepara", "blora", "boyolali", "wonogiri", "kudus", "temanggung", "wonosobo", "sragen", "purbalingga", "purworejo", "rembang", "kabrembang", "kabjepara", "kabboyolali", "kabcilacap", "kabbanyumas", "kabpurbalingga", "rembamg", "kabrembamg", "pemalamg", "purbalimgga", "temamggung", "kabpati", "kabgrobogan", "kabblora" , "bajarnegara", "bamjarnegara", "kabkudus", "kabtemanggung", "kabpemalang", "bamyumas", "cicacap", "kabsragen")

javanese_df$NATIVE <- apply(javanese_df[,2], 1, identify_pob) # apply function to map place of birth, javanese_df[,2] refers to place of birth

javanese_native_df <- javanese_df %>% # exclude those who born not in Central Java
  filter(NATIVE != "migrant")

javanese_native_df

```

## Identifying year of birth (YOB)
It uses National ID Card numbers to identify people's year of birth.

```{r}

##add year of birth
javanese_native_df <- left_join(javanese_native_df, central_java_df[,c(5,9)], by="ID") # 5 refers to NIK, 9 refers to ID
javanese_native_df$NIK <- apply(javanese_native_df[,4], 2, FUN=str_sub, start=-6, end=-5) # 4 refers to NIK
names(javanese_native_df)[names(javanese_native_df) == "NIK"] <- "YOB" # rename NIK with YOB
javanese_native_df$YOB <- as.integer(javanese_native_df$YOB)

#get rid of errors / no YOB
javanese_native_df <- na.omit(javanese_native_df)
which(is.na(javanese_native_df))

#make four digits YOB
javanese_native_df$YOB <- ifelse(javanese_native_df$YOB < 2, javanese_native_df$YOB+2000, javanese_native_df$YOB+1900)

#focus those born in 1945-2001
javanese_native_df <- javanese_native_df %>%
  filter(YOB >= 1945)

```

## Identifying gender
Genders are derived from people's sex as listed in their national ID cards.

```{r}

javanese_native_df <- left_join(javanese_native_df, central_java_df[,8:9], by="ID") # 8 refers to sex, 9 to ID
names(javanese_native_df)[names(javanese_native_df) == "JK"] <- "GENDER" # change column's name JK to Sex
javanese_native_df$GENDER <- recode_factor(javanese_native_df$GENDER, "l" = "male", "p" = "female") # change l to make, p to female

```


# CLEANING 
## Remove symbols

```{r}

javanese_native_df <- left_join(javanese_native_df, central_java_df[,c(6,9)], by="ID") # add names, 6 refers to names, 9 to ID
names(javanese_native_df)[names(javanese_native_df) == "NAMA"] <- "NAME" # change column name 

javanese_native_df$NAME <- str_replace_all(javanese_native_df$NAME, '\\.|,', " ") #get rid of periods and commas

```


## Remove titles

get rid of all acquired titles
academics: se, s pd, sh, st, spd, drs, dra, amd, a md, ba, mm, s ag, dr, sp, s pd i, bsc, msi, s psi, ss, mt, s sos, s h, mh, s ip, ssi
religious and cultural: h, hj, ny

```{r}

acqtitles <- "\\bh\\b|\\bse\\b|\\bhj\\b|\\bs pd\\b|\\bny\\b|\\bsh\\b|\\bst\\b|\\bspd\\b|\\bdrs\\b|\\bdra\\b|\\bamd\\b|\\ba md\\b|\\bba\\b |\\bmm\\b|\\bs ag\\b|\\bdr\\b|\\bsp\\b|\\bs pd i\\b|\\bbsc\\b|\\bmsi\\b|\\bs psi\\b|\\bss\\b|\\bmt\\b|\\bs sos\\b|\\bs h\\b|\\bs ip\\b|\\bssi\\b|\\bdrg\\b|\\bmh\\b"

javanese_native_df$CLEANNAME <- str_remove_all(javanese_native_df$NAME, acqtitles)

```



# COUNTING
## Count lenghts of words

```{r}

javanese_native_df$NAMELEN <- apply(javanese_native_df[,7], 1, nwords) #7 refers to clean name

# get rid of NAs to avoid errors 
which(is.na(javanese_native_df))
javanese_native_df <- na.omit(javanese_native_df)
which(is.na(javanese_native_df))

```

## Count lenghts of characters

```{r}

javanese_native_df$CHARLEN <- apply(javanese_native_df[,7], 1, ncharacters) #7 refers to clean name

# get rid of NAs to avoid errors 
which(is.na(javanese_native_df))
javanese_native_df <- na.omit(javanese_native_df)
which(is.na(javanese_native_df))

```


## Calculate PCA scores

```{r}

# PCA on all
javanese_native_df$PCA <- princomp(javanese_native_df[,8:9], cor=T, scores=T)$scores[,1] #8, 9 referst to name length and characters lenth

which(is.na(javanese_native_df))
javanese_native_df

```

PCA summary

```{r}

pca_namelenght <- princomp(javanese_native_df[,8:9], cor=T, scores=T)
summary(pca_namelenght)
pca_namelenght$loadings

```



# DESCRIPTIVE STATS
This section requires javanese_native_df data set.
## Tables

### Numbers of Javanese people

for manuscript

```{r}

tab_javanese_people <- javanese_native_df %>% 
  mutate(YOB = cut(YOB, breaks = 7)) %>%
  group_by(NATIVE, YOB, GENDER) %>%
  summarise(N_PEOPLE = length(ID)) %>%
  pivot_wider(names_from = c(GENDER, NATIVE), values_from = N_PEOPLE)

tab_javanese_people

```

for supplementary material 1

```{r}

supp_1 <- javanese_native_df %>% 
  group_by(NATIVE, YOB, GENDER) %>%
  summarise(N_PEOPLE = length(ID)) %>%
  pivot_wider(names_from = c(GENDER, NATIVE), values_from = N_PEOPLE)

supp_1

```

### Length of names

for manuscript:

with YOB, POB, and Gender (3 columns)

```{r}

tab_javanese_names_chars_3col <- javanese_native_df %>% 
  group_by(YOB, NATIVE, GENDER) %>%
  summarise(LEN_NAMES = round(mean(NAMELEN, na.rm=T), 2), SD_NAMES = round(sd(NAMELEN, na.rm = T), 2), SE_NAMES = round(stn_err(NAMELEN), 2),
            LEN_CHARS = round(mean(CHARLEN, na.rm=T), 2), SD_CHARS = round(sd(CHARLEN, na.rm = T), 2), SE_CHARS = round(stn_err(CHARLEN), 2))

tab_javanese_names_chars_3col

```

for supplementary material 2

```{r}

supp_2 <- tab_javanese_names_chars_3col %>% 
pivot_wider(names_from = c(GENDER, NATIVE), values_from = LEN_NAMES:SE_CHARS)

supp_2

```

## Visualization

### Length of names with sd

```{r, warning=F}

tab_fig_name_char3 <- tab_javanese_names_chars_3col %>%
  pivot_longer(c(LEN_NAMES, LEN_CHARS), names_to = "Names", values_to = "LENGTH")

tab_fig_name_char3$Names <- recode_factor(tab_fig_name_char3$Names, "LEN_CHARS" = "Number of Characters", "LEN_NAMES" = "Number of Words")
tab_fig_name_char3$NATIVE <- recode_factor(tab_fig_name_char3$NATIVE, "urban" = "Urban", "suburb" = "Suburb", "rural" = "Rural")

tab_fig_name_char3$SD <- as.double(NA)

tab_fig_name_char3[tab_fig_name_char3$Names == "Number of Words", ]$SD <- tab_fig_name_char3[tab_fig_name_char3$Names == "Number of Words", ]$SD_NAMES
tab_fig_name_char3[tab_fig_name_char3$Names == "Number of Characters", ]$SD <- tab_fig_name_char3[tab_fig_name_char3$Names == "Number of Characters", ]$SD_CHARS

#neworder <- c("numbers of names","numbers of characters")
#tab_fig_name_char2 <- arrange(transform(tab_fig_name_char2, Names=factor(Names,levels=neworder)),Names)

fig_name_char_sd_col3 <- tab_fig_name_char3 %>%
  ggplot(aes(x = YOB, y = LENGTH, fill = GENDER, color = GENDER)) +
  geom_ribbon(aes(ymin = (LENGTH - SD), ymax = (LENGTH + SD)), linetype = 0, alpha = 0.17, show.legend = F) +
  geom_line(aes(y = LENGTH), size = 1.5) + scale_x_discrete(name = "Year of Birth", limits = seq(1945, 2001, 7)) + ylab("Mean") + 
  labs(title="The Length of Javanese People's Names Sorted by Year of Birth", 
      color="Gender:")  + 
  facet_grid(Names~NATIVE, scales = "free_y") +
  scale_fill_manual(breaks = c("male", "female"), values=c("#0F3460", "#E94560")) + 
  scale_color_manual(breaks = c("male", "female"), values=c("#0F3460", "#E94560")) + 
  theme(axis.text.x = element_text(angle = 45), panel.spacing.x = unit(1.5, "lines")) 
# + scale_color_manual(values=c("black", "#999999")) + scale_fill_manual(values=c("darkgrey", "lightgrey"))

fig_name_char_sd_col3 + 
  theme(#plot.title = element_text(hjust = 0.5),
        legend.position = "top")

```

### Length of names with standard error

```{r, warning=F}

tab_fig_name_char3$SE <- as.double(NA)

tab_fig_name_char3[tab_fig_name_char3$Names == "Number of Words", ]$SE <- tab_fig_name_char3[tab_fig_name_char3$Names == "Number of Words", ]$SE_NAMES
tab_fig_name_char3[tab_fig_name_char3$Names == "Number of Characters", ]$SE <- tab_fig_name_char3[tab_fig_name_char3$Names == "Number of Characters", ]$SE_CHARS

#neworder <- c("numbers of names","numbers of characters")
#tab_fig_name_char2 <- arrange(transform(tab_fig_name_char2, Names=factor(Names,levels=neworder)),Names)

fig_name_char_se_col <- tab_fig_name_char3 %>%
  ggplot(aes(x = YOB, y = LENGTH, fill = GENDER, color = GENDER)) +
  geom_ribbon(aes(ymin = (LENGTH - SE), ymax = (LENGTH + SE)), linetype = 0, alpha = 0.17, show.legend = F) +
  geom_line(aes(y = LENGTH), size = 1.5) + scale_x_discrete(name = "Year of Birth", limits = seq(1945, 2001, 7)) + ylab("Mean") + 
  labs(title="The Length of Javanese People's Names Sorted by Year of Birth",
       color="Gender")  + 
  facet_grid(Names~NATIVE, scales = "free_y") + #+ theme(axis.text.x = element_text(angle = 90)) # + scale_color_manual(values=c("black", "#999999")) + scale_fill_manual(values=c("darkgrey", "lightgrey"))
  scale_fill_manual(breaks = c("male", "female"), values=c("#0F3460", "#E94560")) + 
  scale_color_manual(breaks = c("male", "female"), values=c("#0F3460", "#E94560")) + 
  theme(axis.text.x = element_text(angle = 45), panel.spacing.x = unit(1.5, "lines")) 
  
fig_name_char_se_col +
    theme(#plot.title = element_text(hjust = 0.5),
        legend.position = "top")

```


# INFERENTIAL STATS
## T-test on PCA score 
ttest_results_PCA
ttest_results_list_PCA
theshold for significant: .001



```{r}

# create a tibble for t-test results
YOB <- c(seq(1945,2001)) #year of birth
POB <- c("urban", "suburb", "rural") #place of birth
ATV <- NA #t statistics 
PVL <- NA #p value for gender
ttest_results_PCA <- tibble(expand.grid(YOB, POB), ATV, PVL)
colnames(ttest_results_PCA) <- c("YOB", "POB", "ATV", "PVL")

# helper
# create a vector to summarize t-test results
ttest_results_list_PCA <- vector(mode="list", length = nrow(ttest_results_PCA))
# for loops for t-test
for(i in 1:nrow(ttest_results_PCA)) {
  ttest_results_list_PCA[[i]] <- t.test(PCA ~ GENDER, data=javanese_native_df[javanese_native_df$YOB==ttest_results_PCA$YOB[i] & javanese_native_df$NATIVE==ttest_results_PCA$POB[i], ])
}

# put values back to the t-test tibble
for(i in 1:nrow(ttest_results_PCA)) {
  ttest_results_PCA$ATV[i] <- ttest_results_list_PCA[[i]]$statistic #no absolute, t statistics
  ttest_results_PCA$PVL[i] <- round(ttest_results_list_PCA[[i]]$p.value, 5)
}

# add significant levels
ttest_results_PCA$SIG <- NA
for(i in 1:nrow(ttest_results_PCA)) {
  if (ttest_results_PCA$PVL[i] >= .001) {
    ttest_results_PCA$SIG[i] <- "no"
  }
  else {
    ttest_results_PCA$SIG[i] <- "yes"
  }
}

ttest_results_PCA

```

for supplementary material 3

```{r}

supp_3 <- ttest_results_PCA %>% 
  select(-SIG) %>% 
  pivot_wider(names_from = c(POB), values_from = ATV:PVL)

supp_3

```

## Viz T-test on PCA (all)
ttest_results__PCA 

```{r, warning=F}

nosig_x <- as.list(ttest_results_PCA[ttest_results_PCA$SIG=="no", 1]) #1 refers to YOB
nosig_y <- as.list(ttest_results_PCA[ttest_results_PCA$SIG=="no", 3]) #1 refers to ATV
nosig_z <- as.list(ttest_results_PCA[ttest_results_PCA$SIG=="no", 2]) #2 refers to POB 
ann_dot <- data.frame(nosig_x, nosig_y, nosig_z)

POB_labels <- c("Urban", "Suburb", "Rural" )
names(POB_labels) <- c("urban", "suburb", "rural")

fig_ttest_results_PCA_col <- ttest_results_PCA %>%
  ggplot(aes(x=YOB, y=ATV)) + geom_line(size=1.2, color="#D55E00") + scale_x_discrete(name="Year of Birth", limits=seq(1945, 2001, 7)) + ylab("T-Statistics") + 
  ggtitle("Gender differences in Javanese people's full names") + 
  labs(caption=" \u2022 = not significant, p > .001") + geom_point(data=ann_dot, color="black") + 
  facet_wrap(~POB,
             labeller = labeller(POB = POB_labels)) + 
  theme(panel.spacing.x = unit(1.2, "lines")) + 
  theme(axis.text.x = element_text(angle = 45)) +
  theme(plot.margin = unit(c(0.2,.5,0.2,0.2), "cm")) #t,r, b, l

fig_ttest_results_PCA_col

```

## Cohen's D effect size on PCA score

```{r}

# create a tibble for Cohen's D effect size results
YOB <- c(seq(1945,2001)) #year of birth
POB <- c("urban", "suburb", "rural") #place of birth
EFS <- NA #effect size
MAG <- NA #magnitude
cohenD_results_PCA <- tibble(expand.grid(YOB, POB), EFS, MAG)
colnames(cohenD_results_PCA) <- c("YOB", "POB", "EFS", "MAG")

# helper
# create a vector to summarize effect size results
cohenD_results_list_PCA <- vector(mode="list", length = nrow(cohenD_results_PCA))
# for loops for t-test
for(i in 1:nrow(cohenD_results_PCA)) {
  cohenD_results_list_PCA[[i]] <- cohens_d(PCA ~ GENDER, data=javanese_native_df[javanese_native_df$YOB == cohenD_results_PCA$YOB[i] & javanese_native_df$NATIVE == cohenD_results_PCA$POB[i], ])
}

# put values back to the Cohen's D tibble
for(i in 1:nrow(cohenD_results_PCA)) {
  cohenD_results_PCA$EFS[i] <- round(cohenD_results_list_PCA[[i]]$effsize,3) #no absolute
  cohenD_results_PCA$MAG[i] <- cohenD_results_list_PCA[[i]]$magnitude
}

cohenD_results_PCA

```
for supplementary material 4

```{r}

supp_4 <- cohenD_results_PCA %>% 
  select(-MAG) %>% 
  pivot_wider(names_from = c(POB), values_from = EFS)

supp_4

```

## Viz Cohen's D on PCA score

```{r}

cohenD_results_PCA %>%
  ggplot(aes(x=YOB, y=EFS)) + geom_line(size=1.2, color="#16697A") + scale_x_discrete(name="Year of Birth", limits=seq(1945, 2001, 7)) + ylab("Effect Size") + 
  ggtitle("Gender differences in Javanese people's full names") + 
  labs(caption=".00 = negligible, .20 = small, .50 = moderate, .80 = large") + 
  facet_wrap(~POB,
             labeller = (labeller(POB = POB_labels))) + theme(panel.spacing.x = unit(1.5, "lines")) + 
  theme(axis.text.x = element_text(angle = 45)) + ylim(-.15, 1) +
  geom_hline(yintercept = c(.2, .5, .8), color=c("darkgrey")) +
  theme(plot.margin = unit(c(0.2,.5,0.2,0.2), "cm")) #t,r, b, l

```
